Efficient Implementation of the Room Simulator \\ for Training Deep Neural Network Acoustic Models

### 2018-06-17 room_simulator_comp_reduction_interspeech_2018/doc/main.tex
Uploads INTERSPEECH 2018 paper

% Template for ASRU-2017 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{amssymb,amsmath,bm}
\usepackage{textcomp}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{multirow}
\usepackage[justification=centering]{caption}
\usepackage[caption = false]{subfig}

@@

\graphicspath{{../figures/}}

@@

% Title.
% ------
\title{Efficient Implementation of the Room Simulator \\ for Training Deep Neural Network Acoustic Models}
%
% Single address.
% ---------------
\name{{Chanwoo Kim, Ehsan Variani, Arun Narayanan, and Michiel Bacchiani}}

\address{Google Speech \\
%  {\small \tt \{chanwcom, tsainath, arunnt, misra, rnongpiur, michiel\}@google.com }}
  {\small \tt \{chanwcom, variani, arunnt, michiel\}@google.com }}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
%
\maketitle
%
\begin{abstract}
In this paper, we describe how to efficiently implement
an acoustic room simulator to generate large-scale
simulated data for training deep neural networks.
Even though the \textit{Room Simulator} in \cite{C_Kim_INTERSPEECH_2017_1} was shown
 to be quite effective in reducing the Word
Error Rates (WERs) for far-field applications by generating
simulated far-field training sets, it requires
high dimensional Fast Fourier Transforms (FFTs).
\textit{Room Simulator} in \cite{C_Kim_INTERSPEECH_2017_1} used 
  approximately 80 percent of CPU usage in our CPU/GPU
training architecture. In this work, we implement an efficient OverLap
  Addition (OLA) based filtering using the open-source \texttt{FFTW3}
library. Further, we investigate the effects of the Room
Impulse Response (RIR) lengths. Experimentally, we conclude that we can cut
the tail portions of RIRs whose power is less than 20 \textit{dB}
below the maximum power without sacrificing the speech recognition accuracy.
Using these approaches, we were able to reduce CPU usage for the
room simulator portion down to 9.69 \%
in CPU/GPU training architecture. Profiling result shows that
we obtain 22.4 times speed-up on a single machine and 37.3 times
  speed up on Google's distributed training infrastructure.
\AR{I don't think we should use borg to describe clusters since it's not something known externally.}
 \end{abstract}
%
\begin{keywords}
Simulated data, room acoustics, robust speech recognition, deep learning
\end{keywords}
%
%
\section{Introduction}
With advancements in deep learning,
\cite{Seltzer2013DNNAurora4, Yu2013FeatureLearningDNN, V_Vanhoucke_Deep_Learning_NIPS_Workshop_2011,
G_Hinton_IEEE_Signal_Process_Mag_2012,
T_Sainath_IEEETran_2017_1, T_Sainath_Book_Chapter_2017_1},
speech recognition accuracy has improved dramatically.
Now, speech recognition systems
are used not only on portable devices
but also on standalone devices for far-field speech recognition.
Examples include voice assistant systems such as Amazon Alexa
and Google Home \cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}.
In far-field speech recognition, the impact of noise and reverberation
is much larger than near-field cases. Traditional approaches to far-field
speech recognition include noise robust feature extraction algorithms
\cite{C_Kim_IEEETran_2016_1, U_H_Yapanel_SpeechComm_2008},
on-set enhancement algorithms
\cite{C_Kim_INTERSPEECH_2010_2, C_Kim_INTERSPEECH_2014_2} \AR{Classification seems too narrow!}.
\AR{Maybe change: Recently, we observed that -> Several studies have now shown that} Recently, we observed that training with noisy data generated
by a \textit{Room Simulator} \cite{C_Kim_INTERSPEECH_2017_1}
improves speech recognition accuracy dramatically.
This system has been successfully employed
for training acoustic models for Google Home or Google voice
search \cite{C_Kim_INTERSPEECH_2017_1}.

\textit{Room Simulator} creates millions of virtual rooms
with different dimensions and different number of sound sources
at different locations and Signal-to-Noise Ratios (SNRs).
For every new utterance in the training set, we use a randomly
sampled room configuration, so that the same utterance
is simulated under different acoustic environments in every epoch during training.
As will be seen in Sec. \ref{sec:experimental_results}, if
we generate the simulated utterance only once for each
input example, the performance is worse.
Since different RIRs are applied for the same
utterance at every epoch, the intermediate results cannot be cached.
Therefore, the noisification process requires several large convolution operations. As will be seen
in Sec. \ref{sec:experimental_results}, during training, it uses up to 80 \%
of the whole CPU usage if we use the CPU/GPU architecture
shown in Fig. \ref{fig:gpu_cpu_structure}.

In this paper, we describe our approach to reduce the
computational cost of the \textit{Room Simulator}. We
first switch from \texttt{Kiss FFT} in
\texttt{Eigen3} to Real FFT in \texttt{FFTW3}. Then we
investigate the required length of RIRs for robust
speech recognition.
%As will be seen in Sec.
%\ref{sec:experimental_results}, we observe that we can remove
%the tail portion or an RIR whose power is 20-dB below
%the maximum power of the original RIR.
%This approach reduced the RIR length to 47 \% compared to the
%original RIR length.
%Combined with OLA filtering, on a local machine, the speed-up
%was 25.2 times. On Google Borg, the speed up was 37.3 times.
%
%
%
%
%
\begin{figure}%[!tbp]
  \begin{center}
    \resizebox{85mm}{!}{\input{../figures/gpu_cpu_structure.tex}}
      \caption {
        \label{fig:gpu_cpu_structure}
        Training architecture using cluster of CPUs and GPUs.
        \cite{E_Variani_INTERSPEECH_2017_01}.
        The ``Room Simulator" is on the CPU side to generate
        simulated utterances using room configurations.
        \cite{C_Kim_INTERSPEECH_2017_1}.}
  \end{center}
\vspace{-10mm}
\end{figure}
%
%
%
%
\section{Efficient Implementation of the Room Simulation System}
  \label{sec:implementation}
%
Fig. \ref{fig:gpu_cpu_structure} shows a high level block diagram of
the acoustic model training infrastructure
\cite{E_Variani_INTERSPEECH_2017_01}. Since it is not easy to
run all the front-end processing blocks, such as feature
extraction and Voice Activity Detection (VAD) on Graphics
Processing Units (GPUs), during training, they run on Central Processing
Units (CPUs). Even though Fast Fourier Transform (FFT) of the
\textit{Room Simulator} may be efficiently implemented on GPUs,
since
\textit{Room Simulator} must feed the simulated utterances
to the rest of CPU optimized front-end components, the Room Simulator
is also running on CPUs.
%
%As will be seen in Table \ref{tbl:cpu_percentage}, the ``\texttt{Room Simulator}"
%system described in \cite{C_Kim_INTERSPEECH_2017_1} consumes 80 percent of the
%CPU cycles in the architecture shown in Fig. \ref{fig:gpu_cpu_structure}.
%
%
%
\subsection{Review of the Room Simulator for Data Augmentation}
%
%
\begin{figure}
  \begin{center}
  \subfloat[\label{fig:room_diagram}] {
    \includegraphics[width=70mm]{room_diagram.eps}} \\
  \subfloat[\label{fig:image_method}] {
    \includegraphics[width=70mm]{image_method.eps}}
   \caption{
    (a) A simulated room: There may be multiple microphones, a single target
    sound source, multiple noise sources in a cuboid-shape room with
    acoustically reflective walls. (b) A diagram showing the location
    of the real and the virtual sound sources assuming that the walls
    reflects acoustic wave like a mirror \cite{C_Kim_INTERSPEECH_2017_1}.}
  \end{center}
  \vspace{-10mm}
\end{figure}
%
%
In this section, we briefly review the structure of the Google
Room Simulator for generating simulated utterances
to train acoustic models for speech recognition systems
\cite{C_Kim_INTERSPEECH_2017_1}. We assume a room of a
rectangular cuboid-shape as shown in Fig. \ref{fig:room_diagram}.
Assuming that all the walls of a room reflect acoustically uniformly,
we use the image method to model the Room Impulse Responses (RIRs)
\cite{J_Allen_JASA_1979, E_A_Lehmann_ASPAA_2007, S_G_McGovern_RIR}.

Fig. \ref{fig:image_method} shows how virtual sources
appearing as empty circles are constructed in the image method.
Even though the virtual rooms appear to be created in two-dimensions
in Fig. \ref{fig:image_method}, in the ``\emph{Room Simulator}",
they are constructed in three-dimensions. For example, in our
work for training the acoustic model for Google Home, we consider
$17 \times 17 \times 17 = 4913$ virtual rooms for RIR calculation
\cite{C_Kim_INTERSPEECH_2017_1}.
Following the image method, the impulse response is calculated
using the following equation
\cite{ J_Allen_JASA_1979, E_A_Lehmann_ASPAA_2007}:
\begin{align}
    h[n] = \sum_{i = 0}^{I-1} \frac{r^{g_i}}{d_i}
    \delta \left[n -\left \lceil{\frac{d_i f_s}{c_0}}\right \rceil \right],
      \label{eq:h_n_calculation}
\end{align}
where $i$ is the index of each virtual sound source, and $d_i$ is the distance
from that sound source to the microphone, $r$ is
the reflection coefficient of the wall, $g_i$ is
the number of the reflections
to that sound source, $f_s$ is the sampling rate of the RIR, and $c_0$ is the
speed of sound in the air. We use the value of $f_s = 16,000$ \textit{Hz} and
$c_0 = 343$ \textit{m/s} for the room simulator \cite{B_Li_INTERSPEECH_2017_1}.
For $d_i$, $r$, we use numbers created by a random number generator
following specified distributions for each impulse response \cite{C_Kim_INTERSPEECH_2017_1}.

Assuming that there are $I$ sound sources including one target
source and $J$ microphones, the received signal at microphone
$j$ is given by:
\begin{align}
  y_j[n] =  \sum_{i=0}^{I-1} \alpha_{ij} \left(h_{ij}[n] * x_i[n]\right).
  \label{eq:y_j_def}
\end{align}
Since we used a two-microphones system in
\cite{B_Li_INTERSPEECH_2017_1, C_Kim_INTERSPEECH_2017_1},
$J$ is two, and for the number noise sources, we used a value from zero
up to three \cite{C_Kim_INTERSPEECH_2017_1}.
%
%
%
%
%
%
%\begin{figure}[!tbp]
%  \begin{center}
%    \resizebox{75mm}{!}{\input{../figures/entire_diagram.tex}}
%      \caption {
%        The architecture for acoustic model training using
%        the room simulator and LSTMs and a DNN (LDNN)
%     \cite{T_Sainath_ICASSP_2016_1, C_Kim_INTERSPEECH_2017_1}. }
%          \label{fig:entire_diagram}
%  \end{center}
%\vspace{-7mm}
%\end{figure}
%
%
%
%%
%\begin{figure}%[tbp]
%  \begin{center}
%    \includegraphics[width=70mm]{reverb_time_distribution}
%    \label{fig:target_length}
%    \caption {
%      Reverberation time ($T_{60}$) distribution in our
%      configuration for training the acoustic model for Google Home
%      and the real Google Home user usage.
%    }
%   \vspace{-7mm}
%  \end{center}
%\end{figure}
%

%
%
%
\subsection{Efficient Room Impulse Response filtering}
%
%
%
When the signal and the room impulse response length are $N_x$ and
$M$, respectively, to calculate \eqref{eq:y_j_def} using the time-domain
convolution, the number of multiplications $C_{td}$ is given by:
\begin{align}
  C_{td} = I \times J \times  M  \times N_x.
  \label{eq:num_mults_linear_convolution}
\end{align}
In our training set used in Sec. \ref{sec:experimental_results},
the average utterance length including \emph{non-speech} portions
marked by the Voice Activity Detector (VAD) is 6.35 \textit{s}.
This corresponds to $N_x$ of 101,523 at 16 \textit{kHz}. The average
length of the RIR in the room simulator used in
\cite{C_Kim_INTERSPEECH_2017_1} is 0.243 \textit{s} \AR{Sounds a bit too smalle?}, corresponding
to the $M$ value of 3893.
For Google Home the number of noise sources in our training
set is 1.55 on average.
Thus $I$ is 1.55 including one target source, and $J$ is two, since
we use a two-microphones system in
\cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}.
Thus, if we directly use the linear convolution, it requires
two billion multiplications per utterance from
\eqref{eq:num_mults_linear_convolution}, which is prohibitively large.
Thus, in the ``\textit{Room Simulator}" in
\cite{C_Kim_INTERSPEECH_2017_1},
we used the frequency domain multiplication using \texttt{Kiss FFT}
\cite{M_Borgerding_kiss_fft_2010}. To avoid time aliasing,
the FFT size $N$ must satisfy $N \ge N_x + M - 1$.
The $j$-th microphone channel
of the simulated signal $y_j[n]$ in \eqref{eq:y_j_def}
\begin{align}
  y_{j} [n] = \sum_{i=0}^{I-1}FFT^{-1} \left \{
      FFT \left \{ h_{ij}[n] \right \} \times
      FFT \left \{ x_i[n] \right \}
      \right \}  %* FFT\{X_i[omega_k}\}\}
\end{align}
Considering we perform two FFTs \AR{Why 2? You should consider speech and 1.55 noises right?}, one IFFT, and one complex spectrum
element-wise multiplications, the number of real multiplications
$C_{\textit{FFT}}$ is given by:
\begin{align}
  C_{\textit{FFT}} = 6 N \log_2(N) + 4 N.
  \label{eq:c_fft}
\end{align}
where $N$ is the FFT size. \AR{Shouldn't all this be in big-O notation? So not exactly NlogN but O(NlogN)?} The number of real multiplications
in \eqref{eq:c_fft} is based on the assumption that a single FFT 
requires $\frac{N}{2} \log_2(N)$ multiplications \AR{This is confusing, I think it makes sense to just use NlogN given you are already ignoring other constants that are not O(NlogN).}, and a single
complex multiplication requires four real multiplications. The $4N$
term in \eqref{eq:c_fft} appears since there is one frequency domain
element-wise multiplications.
If we use $N = 2 ^ {17}$, it requires 70.9 million multiplications
per utterance. The room simulator was implemented in C++ using
the \texttt{Eigen3} linear algebra library \cite{eigenweb}.
In \texttt{Eigen3}, \texttt{Kiss FFT} \cite{M_Borgerding_kiss_fft_2010} and \texttt{FFTW3}
\cite{M_Frigo_proceedings_of_the_ieee_20015} are unofficially supported.
We used the \texttt{Kiss FFT} version of FFT in \texttt{Eigen3} since
there were some technical issues in using \texttt{FFTW3} in \texttt{Eigen3}
to run on \texttt{Google Borg} cluster \cite{A_Verma_eurosys_2015_1}.
But, to fruther speed up frontend computation, we switched from
\texttt{Kiss FFT} in \texttt{Eigen3} to a custom  C++ class
implementation which internally using real FFT in \texttt{FFTW3}.
\AR{The argument seems weak. Consider rephrasing? Maybe say, our initial implementation was based on KissFFT. But, to further speed up computation, we switched ... i.e., don't mention about technical issues on borg cluster -- that seems something very specific and scientifically uninteresting.}

>>>>

% Template for ASRU-2017 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass[a4paper]{article}

\usepackage{INTERSPEECH2016}

\usepackage{amsmath,graphicx}
\usepackage{amssymb,amsmath,bm}
\usepackage{textcomp}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{multirow}
\usepackage[justification=centering]{caption}
\usepackage[caption = false]{subfig}

@@

\graphicspath{{.}{../figures/}}
\makeatletter
  \def\input@path{{.}{../figures/}}
\makeatother

@@

% Title.
% ------
\title{Efficient Implementation of the Room Simulator \\ for Training Deep Neural Network Acoustic Models}
%
% Single address.
% ---------------
\name{{Chanwoo Kim$^{1\dagger}$\thanks{$^\dagger$Work performed while at Google.}, Ehsan Variani$^2$, Arun Narayanan$^2$, and Michiel Bacchiani$^2$}}

\address{$^1$Samsung Research, $^2$Google Speech \\
%  {\small \tt \{chanwcom, tsainath, arunnt, misra, rnongpiur, michiel\}@google.com }}
  {\small \tt $^1$chanw.com@samsung.com $^2$\{variani, arunnt, michiel\}@google.com }}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
%
\maketitle
%
\begin{abstract}

In this paper, we describe how to efficiently implement
an acoustic room simulator to generate large-scale
simulated data for training deep neural networks.
Even though \textit{Google Room Simulator} in \cite{C_Kim_INTERSPEECH_2017_1} was shown
 to be quite effective in reducing the Word
Error Rates (WERs) for far-field applications by generating
simulated far-field training sets, it requires a very large 
number of FFTs.
\textit{Room Simulator} used approximately 80 \% of CPU usage
  in our \textit{CPU/GPU} training architecture
  \cite{E_Variani_INTERSPEECH_2017_01}.
  In this work, we implement an efficient OverLap
  Addition (OLA) based filtering using the open-source \texttt{FFTW3}
library. Further, we investigate the effects of the Room
Impulse Response (RIR) lengths. Experimentally, we conclude that we can cut
the tail portions of RIRs whose power is less than 20 \textit{dB}
below the maximum power without sacrificing the speech recognition accuracy.
However, we observe that cutting  RIR tail more than this threshold
harms the speech recognition accuracy for rerecorded test sets.
Using these approaches, we were able to reduce CPU usage for the
room simulator portion down to 9.69 \%
in CPU/GPU training architecture. Profiling result shows that
we obtain 22.4 times speed-up on a single machine and 37.3 times
  speed up on Google's distributed training infrastructure.
 \end{abstract}
%
%\begin{keywords}
  \noindent{\bf Index Terms}: Simulated data, room acoustics, robust speech recognition, deep learning
%\end{keywords}
%
%
\section{Introduction}
With advancements in deep learning
\cite{Seltzer2013DNNAurora4, Yu2013FeatureLearningDNN, V_Vanhoucke_Deep_Learning_NIPS_Workshop_2011,
G_Hinton_IEEE_Signal_Process_Mag_2012,
T_Sainath_IEEETran_2017_1, T_Sainath_Book_Chapter_2017_1},
speech recognition accuracy has improved dramatically.
Now, speech recognition systems
are used not only on portable devices
but also on standalone devices for far-field speech recognition.
Examples include voice assistant systems such as Amazon Alexa
and Google Home \cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}.
In far-field speech recognition, the impact of noise and reverberation
is much larger than near-field cases. Traditional approaches to far-field
speech recognition include noise robust feature extraction algorithms
\cite{C_Kim_IEEETran_2016_1, U_H_Yapanel_SpeechComm_2008, C_Kim_ICASSP_2010_1},
on-set enhancement algorithms
\cite{C_Kim_INTERSPEECH_2010_2, C_Kim_INTERSPEECH_2014_2}, and multi-microphone
approaches \cite{T_Nekatani_ICASSP_2017_1, T_Higuchi_ICASSP_2016_1,
H_Erdogan_INTERSPEECH_2016_1, C_Kim_INTERSPEECH_2015_1, C_Kim_INTERSPEECH_2009_1, C_Kim_ICASSP_2012_2, C_Kim_ICASSP_2011_2}.
Recently, we observed that training with large-scale noisy data generated
by a \textit{Room Simulator} \cite{C_Kim_INTERSPEECH_2017_1}
improves speech recognition accuracy dramatically.
This system has been successfully employed
for training acoustic models for Google Home or Google voice
search \cite{C_Kim_INTERSPEECH_2017_1}.

\textit{Room Simulator} creates millions of virtual rooms
with different dimensions and different number of sound sources
at different locations and Signal-to-Noise Ratios (SNRs).
For every new utterance in the training set, we use a randomly
sampled room configuration, so that the same utterance
is simulated under different acoustic environments in every epoch during training.
As will be seen in Sec. \ref{sec:experimental_results}, if
we generate the simulated utterance only once for each
input example, the performance is worse.
Since different RIRs are applied for the same
utterance at every epoch, the intermediate results cannot be cached.
Therefore, the noisification process requires very large number of convolution operations. As will be seen
in Sec. \ref{sec:experimental_results}, during training, \textit{Room Simulator} used up to 80 \%
of the whole CPU usage if we use the CPU/GPU architecture
shown in Fig. \ref{fig:gpu_cpu_structure}.
In this paper, we describe our approach to reduce the
computational portion of the \textit{Room Simulator} below 10 \% of the CPU usage
in our CPU/GPU training scheme.
%We
%first switch from \texttt{Kiss FFT} in
%\texttt{Eigen3} to Real FFT in \texttt{FFTW3}. Then we
%investigate the required length of RIRs for robust
%speech recognition.
%As will be seen in Sec.
%\ref{sec:experimental_results}, we observe that we can remove
%the tail portion or an RIR whose power is 20-dB below
%the maximum power of the original RIR.
%This approach reduced the RIR length to 47 \% compared to the
%original RIR length.
%Combined with OLA filtering, on a local machine, the speed-up
%was 25.2 times. On Google Borg, the speed up was 37.3 times.
\begin{figure}%[!tbp]
    \resizebox{75mm}{!}{\input{gpu_cpu_structure.tex}}
        \label{fig:gpu_cpu_structure}
        Training architecture using cluster of CPUs and GPUs.
        \cite{E_Variani_INTERSPEECH_2017_01}.
        \textit{Room Simulator} is on the CPU side to generate
        simulated utterances using room configurations.
        \cite{C_Kim_INTERSPEECH_2017_1}.}
\vspace{-10mm}
\section{Efficient Implementation of the Room Simulation System}
  \label{sec:implementation}
Fig. \ref{fig:gpu_cpu_structure} shows a high level block diagram of
the acoustic model training infrastructure
\cite{E_Variani_INTERSPEECH_2017_01}. Since it is not easy to
run all the front-end processing blocks such as feature
extraction and Voice Activity Detection (VAD) on GPUs,
 they run on CPUs during training.
Even though Fast Fourier Transform (FFT) of the
\textit{Room Simulator} may be efficiently implemented on GPUs,
since
\textit{Room Simulator} must feed the simulated utterances
to the rest of CPU optimized front-end components, the Room Simulator
is also running on CPUs.
%As will be seen in Table \ref{tbl:cpu_percentage}, the ``\texttt{Room Simulator}"
%system described in \cite{C_Kim_INTERSPEECH_2017_1} consumes 80 percent of the
%CPU cycles in the architecture shown in Fig. \ref{fig:gpu_cpu_structure}.
\subsection{Review of the Room Simulator for Data Augmentation}
%  \subfloat[\label{fig:room_diagram}] {
    \includegraphics[width=70mm]{room_diagram.pdf}
% \subfloat[\label{fig:image_method}] {
%   \includegraphics[width=70mm]{image_method.eps}}
   \caption{
    A simulated room: There may be multiple microphones, a single target
    sound source, multiple noise sources in a cuboid-shape room with
    acoustically reflective walls \cite{C_Kim_INTERSPEECH_2017_1}. 
    \label{fig:room_diagram}
    %(b) A diagram showing the location
%    of the real and the virtual sound sources assuming that the walls
%    reflects acoustic wave like a mirror \cite{C_Kim_INTERSPEECH_2017_1}.}
  \vspace{-10mm}
In this section, we briefly review the structure of the Google
Room Simulator for generating simulated utterances
to train acoustic models for speech recognition systems
\cite{C_Kim_INTERSPEECH_2017_1}. We assume a room of a
rectangular cuboid-shape as shown in Fig. \ref{fig:room_diagram}.
Assuming that all the walls of a room reflect acoustically uniformly,
we use the image method to model the Room Impulse Responses (RIRs)
\cite{J_Allen_JASA_1979, E_A_Lehmann_ASPAA_2007,
s_g_mcgovern_mathworks_file_exchange_2013_00,
s_g_mcgovern_applied_acoustics_2009_00}.
In the image method, a real room is acoustically mirrored
with respect to each wall, which results in grids of virtual rooms.
%Fig. \ref{fig:image_method} shows how virtual sources
%appearing as empty circles are constructed in the image method.
%Even though the virtual rooms appear to be created in two-dimensions
%in Fig. \ref{fig:image_method}, in the ``\emph{Room Simulator}",
%they are constructed in three-dimensions. 
In our work for training the acoustic model for Google Home
\cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}, we consider
$17 \times 17 \times 17 = 4913$ virtual rooms for RIR calculation.
Following the image method and assuming that there are $V$ 
virtual sound sources including one real source, 
the impulse response is calculated using the following equation
\cite{ J_Allen_JASA_1979, E_A_Lehmann_ASPAA_2007}:
\begin{align}
    h[n] = \sum_{v = 0}^{V-1} \frac{r^{g_v}}{d_v}
    \delta \left[n -\left \lceil{\frac{d_v f_s}{c_0}}\right \rceil \right],
      \label{eq:h_n_calculation}
\end{align}
where $v$ is the index of each virtual sound source, and $d_v > 0$ is the distance
from that sound source to the microphone, $0 < r < 1$ is
the reflection coefficient of the wall, $g_v$ is
the number of the reflections
to that sound source, $f_s$ is the sampling rate of the RIR, and $c_0$ is the
speed of sound in the air. We use the value of $f_s = 16,000$ \textit{Hz} and
$c_0 = 343$ \textit{m/s} for \textit{Room Simulator} \cite{B_Li_INTERSPEECH_2017_1}.
For $d_v$, $r$, we use numbers created by a random number generator
following specified distributions for each impulse response \cite{C_Kim_INTERSPEECH_2017_1}.

Assuming that there are $I$ sound sources including one target
source and $J$ microphones, the received signal at microphone
$j$ is given by:
\begin{align}
  y_j[n] =  \sum_{i=0}^{I-1} \alpha_{ij} \left(h_{ij}[n] * x_i[n]\right).
  \label{eq:y_j_def}
\end{align}
Since we used a two-microphones system in
\cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}
$J$ is two, and for the number of noise sources, we used a value from zero
up to three with an average of 1.55. Including the target source, the average
number of $I$ is 2.55 \cite{C_Kim_INTERSPEECH_2017_1}.
%
%
%
%
%
%
%\begin{figure}[!tbp]
%  \begin{center}
%    \resizebox{75mm}{!}{\input{../figures/entire_diagram.tex}}
%      \caption {
%        The architecture for acoustic model training using
%        the room simulator and LSTMs and a DNN (LDNN)
%     \cite{T_Sainath_ICASSP_2016_1, C_Kim_INTERSPEECH_2017_1}. }
%          \label{fig:entire_diagram}
%  \end{center}
%\vspace{-7mm}
%\end{figure}
%
%
%
%%
%\begin{figure}%[tbp]
%  \begin{center}
%    \includegraphics[width=70mm]{reverb_time_distribution}
%    \label{fig:target_length}
%    \caption {
%      Reverberation time ($T_{60}$) distribution in our
%      configuration for training the acoustic model for Google Home
%      and the real Google Home user usage.
%    }
%   \vspace{-7mm}
%  \end{center}
%\end{figure}
%

%
%
%
\subsection{Efficient Room Impulse Response filtering}
%
%
%
When the signal and the room impulse response lengths are $N_x$ and
$N_h$ respectively, the number of multiplications $C_{td}$ required for
calculating \eqref{eq:y_j_def} using the time-domain convolution is given by:
\begin{align}
  C_{td} = I \times J \times N_x \times N_h.
  \label{eq:num_mults_linear_convolution}
\end{align}
In our training set used in \cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1},
the average utterance length including \emph{non-speech} portions
marked by the Voice Activity Detector (VAD) is 7.31 \textit{s}.
This corresponds to $N_x$ of 116,991 at 16 \textit{kHz}. The average
length of the RIR in the room simulator used in
\cite{C_Kim_INTERSPEECH_2017_1} is 0.243 \textit{s}, corresponding
to the $N_h$ value of 3893.
For Google Home the number of noise sources in our training
set is 1.55 on average.
Thus $I$ is 2.55 including one target source, and $J$ is two, since
we use a two-microphones system in
\cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}.
Thus, if we directly use the linear convolution, it requires
2.32 billion multiplications per utterance from
\eqref{eq:num_mults_linear_convolution}, which is prohibitively large.
Thus, in the ``\textit{Room Simulator}" in
\cite{C_Kim_INTERSPEECH_2017_1},
we used the frequency domain multiplication using \texttt{Kiss FFT}
\cite{M_Borgerding_kiss_fft_2010}. To avoid time aliasing,
the FFT size $N$ must satisfy $N \ge N_x + M - 1$, where
$M$ is the length of the impulse response.
The $j$-th microphone channel
of the simulated signal $y_j[n]$ in \eqref{eq:y_j_def}
is given by:
\begin{align}
  y_{j} [n] = \sum_{i=0}^{I-1}FFT^{-1} \left \{
      FFT \left \{ h_{ij}[n] \right \} \times
      FFT \left \{ x_i[n] \right \}
      \right \}.  %* FFT\{X_i[omega_k}\}\}
  \label{eq:fir_fft_convolution}
\end{align}
As shown in \eqref{eq:fir_fft_convolution}, we perform
two FFTs, one IFFT, and one complex element-wise multiplications
between two complex spectra for each convolution term in \eqref{eq:y_j_def}.
Assuming that radix-2 FFTs are employed, each FFT or IFFT requires
$\frac{N}{2} \log_2(N)$ multiplications
\cite{A_V_Oppenheim_PrenticeHall_1999}.
In addition, a single element-wise complex multiplication
between two complex spectra is required for each convolution
term in \eqref{eq:y_j_def}.
For real time-domain signals, we need to perform element-wise
multiplications for the lower half-spectrum for
Discrete Fourier Transform (DFT) indices
$0 \le k \le N / 2$ since the spectrum has the Hermitian symmetry
property \cite{A_V_Oppenheim_PrenticeHall_1999}.
From this discussion, we conclude that the number of real multiplications
$C_{\textit{FFT}}$ for calculating \eqref{eq:fir_fft_convolution} is given by:
\begin{align}
  C_{\textit{FFT}} = I J \left(6 N \log_2(N) + 2 N\right).
  \label{eq:c_fft}
\end{align}
For the average $N_x$ of 116,991 mentioned above,
if we assume that the average $N$ is $2 ^ {17}$,  \eqref{eq:c_fft} 
requires 69.5 million multiplications per utterance on average. 
\textit{Room Simulator} for \cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}
was implemented in C++ using the \texttt{Eigen3} linear algebra library \cite{eigenweb}.
%In \texttt{Eigen3}, \texttt{Kiss FFT} \cite{M_Borgerding_kiss_fft_2010} and \texttt{FFTW3}
%\cite{M_Frigo_proceedings_of_the_ieee_20015} are unofficially supported.
So far we have used the \texttt{Kiss FFT} version of FFT in \texttt{Eigen3}
including acoustic model training for Google Home described in
\cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}.
But, to further speed up frontend computation, we switched from
\texttt{Kiss FFT} in \texttt{Eigen3} to a custom  C++ class
implementation which internally uses real FFT in \texttt{FFTW3}.
